---
title: "Open gast_Aug17 set and filter genotypes"
author: "EF"
date: "04 Aug 2017"
output: html_document
---
# start_section
  
```{r start_section}

# Time stamp
Sys.time()

# Clean up
rm(list=ls())
graphics.off()

# Set root working folder
library(knitr)
library(dplyr)
library(stringr)

opts_knit$set(root.dir = "/scratch/medgen/users/eleanor/2015-16/IHCAP/gast_Aug17/")
#setwd("/scratch/medgen/users/eleanor/2015-16/IHCAP/gast_Aug17/")

```

#Load data
```{r load}


vv <- read.table("Source_data/gast_Aug17_relaxed_vqsr_std_sma_ann_VV.txt", sep = "\t", header = TRUE, row.names = 1, stringsAsFactors = FALSE)

gt <- read.table("Source_data/gast_Aug17_relaxed_vqsr_std_sma_ann_GT_add.txt", sep = "\t", header = TRUE, row.names = 1, stringsAsFactors = FALSE)

dp <- read.table("Source_data/gast_Aug17_relaxed_vqsr_std_sma_ann_DP.txt", sep = "\t", header = TRUE, row.names = 1, stringsAsFactors = FALSE)

gq <- read.table("Source_data/gast_Aug17_relaxed_vqsr_std_sma_ann_GQ.txt", sep = "\t", header = TRUE, row.names = 1, stringsAsFactors = FALSE)

exac <- read.table("Source_data/gast_Aug17_relaxed_vqsr_std_sma_ann_exac.txt", sep = "\t", header = TRUE, row.names = 1, stringsAsFactors = FALSE)

kgen <- read.table("Source_data/gast_Aug17_relaxed_vqsr_std_sma_ann_kgen.txt", sep = "\t", header = TRUE, row.names = 1, stringsAsFactors = FALSE)
```

#Convert to matrix
```{r matrix}
#Convert to matrix
vv.mx <- as.matrix(vv)

gt.mx <- as.matrix(gt)

rm(vv, gt)

#check dimensions before filtering
dim(vv.mx)

dim(gt.mx)
```


#Filtering notes
Filters consistent with filtering applied to WC1K. Call rate filter was lowered to < 0.6 as the dataset is much smaller than WC1K and easier to look through manually, plus there may be families with data not missing that is still useful, consider removing this filter completely.
The filters are applied in the following order, filtering out:  
gq < 20  
dp > 500  
call_rate < 0.5

#Filter on gq
```{r gq filter}

#convert to matrix
gq.mx <- as.matrix(gq)

# num of genotypes to be removed
sum(gq.mx < 20, na.rm=TRUE) #346070

# Fraction of genotypes to be removed
sum(gq.mx < 20, na.rm=TRUE)/sum(!is.na(gq.mx)) # ~6.7%

# Apply filter (to gt only !), can only run on matrices
NA -> gt.mx[ gq.mx < 20 ]

rm(gq, gq.mx)
```

#Filter on dp
```{r dp filter}
#convert to matrix
dp.mx <- as.matrix(dp)

# num of genotypes to be removed
sum(dp.mx > 500, na.rm=TRUE) #3776

# Fraction of genotypes to be removed
sum(dp.mx > 500, na.rm=TRUE)/sum(!is.na(dp.mx)) # ~0.0007%

# Apply filter (to gt only !), can only run on matrices
NA -> gt.mx[ dp.mx > 500 ]

rm(dp, dp.mx)
```

#Filter on call rate
```{r callrate filter}

#Create callrates
cr <- data.frame(RawVarID=row.names(gt.mx), callrate = NA)
cr$callrate <- apply(gt.mx, 1, function(x) sum(!is.na(x))/ncol(gt.mx))  

# num of genotypes to be removed
sum(cr$callrate < 0.5) #3620

# Fraction of genotypes to be removed
sum(cr$callrate < 0.5, na.rm=TRUE)/sum(nrow(cr)) # ~2.9%

#Filter variants from cr

cr.filt <- cr[cr$callrate >= 0.5,]

# Remove variants from gt, vv, exac and kgen
gt.filt <- gt.mx[row.names(gt.mx) %in% cr.filt$RawVarID,]

vv.filt <- vv.mx[row.names(vv.mx) %in% cr.filt$RawVarID,]

exac.filt <- exac[row.names(exac) %in% cr.filt$RawVarID,]

kgen.filt <- kgen[row.names(kgen) %in% cr.filt$RawVarID,]

rm(cr, cr.filt, gt.mx, vv.mx, exac, kgen)
```

#Check rownames
```{r check rownames}

sum(row.names(gt.filt) != row.names(vv.filt)) 

sum(row.names(gt.filt) != row.names(exac.filt))

sum(row.names(gt.filt) != row.names(kgen.filt))

```

#Write data
```{r write}
save.image("R_data/s01_gast_Aug17_relax_v2_20170804.Rdata")
```

